### 2014-12-06 docs/ms.Rmd
Mostly tightening the intro. I almost like it.

Studies of biological colouration have generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010]. The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003]. These developments are positive for the field, and have allowed increasing numbers of researchers to rapidly collect and analyse ever greater amounts of data. This rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher. As a consequence, the subtle complexity of many analytical techniques [e.g. the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported.

Replication and transparency lie at the heart of the scientific enterprise. While the full replication of individual studies may often be infeasible [@peng_reproducible_2011], the reproductioin of a study's findings by independent researchers is a readily attainable compromise [@peng_reproducible_2006; @peng_reproducible_2011]. A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved. While completely reproducible research [e.g. @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable. Nevertheless, even partial reproducibility through the simple practice of complete methods reporting and public data archiving is of tremendous value. It ensures greater comparability between studies, allows for the the independent verification of results, and provides a foundation for the testing of new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].

Our aim here to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be improved in line with points (i) - (iii) above. We begin by briefly outlining some of the most common methods for studying biological colouration, and the way in which.... We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the current state of methods reporting in colour science. We also assess the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research.

@@

Although relatively easy to implement, and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually processed, that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012]. For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; -@vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e. raw versus log-transformed), the use of any colour constancy mechanism (e.g. von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e. neural and/or receptor noise). These data are only available for a small number of species, such as humans and honeybees, and so assumptions and extrapolations are necessary when models are applied to other species. Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].

@@

To assess the current state of reproducibility in the field we searched papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_). On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract. We read the titles and abstracts of each paper and included only those papers that used either a spectrometer or camera to quantify colouration. We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking micro-spectrometric measurements of retinal absorbance. To review the large number of papers returned from our search, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our assesment, each paper was read and reassessed by three authors. Any discrepancies between assessor scores were discussed and resolved prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?). We have kept the papers used in our dataset anonymous as our aim is to make a general statement about methods reporting in the field. Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.

@@

_Guidelines for methods reporting_

The simplest and most fundamental step in ensuring research is reproducible is through the comprehensive reporting of methods. The myriad researcher degrees of freedom in the capture and analyses of colour data (as outlined above) heighten this need. Accordingly, we developed a list of essential information pertaining to the capture (Table 1) and analysis (Table 2) of data that should be reported to ensure reproducibility. Given the current state of the field, we focus on spectrometry and photography for the capture of data [as in @kemp_ornamental_2008; @chiao_visualization_2009], and do not consider qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009]. Analytical techniques are diverse [kelber_spectral_2010], so we focus on two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling (Table 2). Extra little bit about vis models in particular. These are essential if such analyses are to be reproduced. These criteria provided that baseline against which to hold our results.

@@

Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used the colour hexagon (`r vismod_type[1]` studies), or tetrahedral colour space (`r vismod_type[4]` studies) approach. Case specific models were used in `r vismod_type[2]` papers. Of visual modelling papers, `r round(propSpec_fin["prop_irrad_type"], digits =2)`% reported the type of irradiance when necessary, `r round(propSpec_fin["prop_vis_mod_sp"], digits = 2)`% the species modelled as a receiver. The background used in visual models was reported `r round(propSpec_fin["prop_vis_mod_bkg"], digits = 2)`% of the time. The type of receptor noise modelled was reported in only `r round(propSpec_fin["prop_vis_mod_noise_type"], digits = 2)`% of the studies. Explicit mention of the type of quantum catch (`r round(propSpec_fin["prop_vis_mod_qcatch"], digits = 2)`% of studies), or whether photoreceptor adaptation (`r round(propSpec_fin["prop_vis_mod_adapt"], digits = 2)`%) was modelled, was not common.

@@

(PARA) None of the included studies linked to any form of code...benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.

>>>>

Studies of biological colouration have generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010]. The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003]. While these developments are positive for the field, the increasingly complex analyses being run on ever greater amounts of data heighten the need for comprehensive methods reporting and proper data management [@alsheikh_public_2011; @nekrutenko_next_2012].

Replication and transparency lie at the heart of the scientific enterprise. Beyond simply allowing the independent verification of results, reproducible research engenders greater comparability between studies and provides a foundation for the testing of new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007]. A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved. While completely reproducible research [e.g. @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable. Nevertheless, even partial reproducibility through the relatively simple practices of complete methods reporting and public data archiving is of tremendous value.

Our aim here to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be improved in line with points (i) - (iii) above. We begin by briefly outlining some of the most common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can dramatically shape results. We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the current state of methods reporting in colour science. We also assess the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research.

@@

Although relatively easy to implement, and potentially powerful, visual models are built on multiple assumptions about the way in which stimuli are visually processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012]. For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; -@vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e. raw versus transformed), the use of any colour constancy mechanism (e.g. von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e. neural and/or receptor noise). These data are only available for a small number of species, such as humans and honeybees, and so assumptions and extrapolations are necessary when models are applied to other species. Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].

@@

To assess the current state of reproducibility in the field we searched papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_). On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract. We read the titles and abstracts of each paper and included only those papers that used either a spectrometer or camera to quantify colouration. We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking micro-spectrometric measurements of retinal absorbance. To review the large number of papers returned from our search, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our assesment, each paper was read and reassessed by three authors. Any discrepancies between assessor scores were discussed and resolved prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found on github (WHAT'S THE DOI?). We have kept the papers used in our dataset anonymous as our aim is to explore the general question of repeatability in the field. Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.

@@

_Guidelines for reporting_

The simplest step in ensuring research is reproducible is the comprehensive reporting of methods. The myriad researcher degrees of freedom in the capture and analyses of colour data (as outlined above) heighten this need. Accordingly, we developed a list of essential information about the capture (Table 1) and analysis (Table 2) of data that should be reported to ensure reproducibility. Given their overwhelming popularity, we focus on spectrometry and photography for data collection [e.g. @kemp_ornamental_2008; @chiao_visualization_2009], and do not consider qualitative methods such as the use of Munsell chips or paint swatches [e.g. @hogstad_gloger_2009].

Analytical techniques are diverse and are being developed rapidly [kelber_spectral_2010]. This rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher. As a consequence, the subtle complexity of many analytical techniques [e.g. the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported. Our guidelines for reporting analytical details (Table 2) cover two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling. These are essential if such analyses are to be reproduced. These criteria provided that baseline against which to hold our results.

@@

Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used the colour hexagon (`r vismod_type[1]` studies), or a tetrahedral colour space (`r vismod_type[4]` studies). Case specific models were used in `r vismod_type[2]` papers. Of visual modelling papers, `r round(propSpec_fin["prop_irrad_type"], digits =2)`% reported the type of irradiance when necessary, `r round(propSpec_fin["prop_vis_mod_sp"], digits = 2)`% the species modelled as a receiver. The background used in visual models was reported `r round(propSpec_fin["prop_vis_mod_bkg"], digits = 2)`% of the time. The type of receptor noise modelled was reported in only `r round(propSpec_fin["prop_vis_mod_noise_type"], digits = 2)`% of the studies. Explicit mention of the type of quantum catch (`r round(propSpec_fin["prop_vis_mod_qcatch"], digits = 2)`% of studies), or whether photoreceptor adaptation (`r round(propSpec_fin["prop_vis_mod_adapt"], digits = 2)`%) was modelled, was not common.

@@

None of the included studies linked to any form of code, which is not surprising given the ongoing popularity of graphical-user-interface based statistical [e.g. SPSS], and colour-analytical [e.g. avicol;] software. Benefits of code over point-and-click, maintain chain of mod for yourself and others, open source shit (R, pavo, things), github, sourcefourge etc. This paper was written openly on github,
